---
title: "Field Operations, Users, and Assignments"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Field Operations, Users, and Assignments}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(SurveySolutionsAPIv2)
library(data.table)

# Load pre-rendered demo data for offline vignette compilation
ext_users <- system.file("extdata", "demo_users.rds", package = "SurveySolutionsAPIv2")
if (file.exists(ext_users)) {
  demo_users <- readRDS(ext_users)
} else {
  demo_users <- UserClass(list(Users = data.frame(UserName = "interviewer01", Role = "Interviewer")))
}

ext_ass <- system.file("extdata", "demo_assignments.rds", package = "SurveySolutionsAPIv2")
if (file.exists(ext_ass)) {
  demo_ass <- readRDS(ext_ass)
} else {
  demo_ass <- assignmentClass(list(Assignments = data.frame(Id = 101, ResponsibleName = "interviewer01")))
}
```

## Introduction

Coordinating large-scale field operations involves creating field staff accounts, provisioning regional workspaces, distributing sample assignments to enumerators, and managing spatial maps.

**`SurveySolutionsAPIv2`** automates these operational workflows, enabling seamless integration with sample registries and field monitoring dashboards.

---

## 1. Workspace Administration

Administrators can partition projects into isolated workspaces (e.g. by region, pilot vs. main survey, or distinct administrative units).

> **Note on Permissions**: Workspace administration operations (`suso_createWorkspace`, `suso_updateWorkspace`, `suso_deleteWorkspace`, `suso_enableWorkspace`, and `suso_assignWorkspace`) require server-level **Administrator (`admin`)** credentials. Regular API user accounts are restricted to their assigned workspace and cannot create, modify, or delete workspaces.

### Creating a New Workspace

```r
# Create a dedicated workspace for a regional survey (requires admin credentials)
suso_createWorkspace(
  new_workspace = "central_region",
  displayName = "Central Region Survey 2025"
)
```

### Updating Workspace Details

```r
# Update display name
suso_updateWorkspace(
  workspace = "central_region",
  displayName = "Central Region (Updated Jan 2025)"
)
```

### Enabling and Disabling Workspaces

When fieldwork for a region concludes, disable the workspace to prevent unintended edits while keeping historical data intact:

```r
# Temporarily disable access
suso_enableWorkspace(workspace = "central_region", enable = FALSE)

# Re-enable access
suso_enableWorkspace(workspace = "central_region", enable = TRUE)
```

---

## 2. Managing Users and Field Staff

Field hierarchies typically consist of Headquarters, Supervisors, and Interviewers.

### Batch Creating Enumerators

Rather than creating dozens of enumerator accounts manually via the web interface, `suso_createUSER()` provisions staff in bulk from a standard data frame:

```r
# Define field staff to create
new_staff <- data.frame(
  Role       = c("Supervisor", "Interviewer", "Interviewer"),
  UserName   = c("supervisor01", "interviewer01", "interviewer02"),
  Password   = c("SecurePass1!", "SecurePass1!", "SecurePass1!"),
  Supervisor = c(NA, "supervisor01", "supervisor01"),
  FullName   = c("Sarah Supervisor", "John Interviewer", "Amina Interviewer"),
  stringsAsFactors = FALSE
)

# Create staff accounts on the server
result <- suso_createUSER(userlist = new_staff)
```

### Querying Field Staff with `UserClass`

To inspect active field staff, retrieve supervisors or interviewers:

```r
# List all supervisors
supervisors <- suso_getSV()

# List all interviewers and supervisors
all_users <- suso_getUSR()
```

The returned object inherits from `UserClass` (and `data.table`), enabling direct table operations and custom summaries:

```{r}
# Display active staff
demo_users[, .(UserName, Role, FullName, Supervisor)]
```

Check class membership:

```{r}
is.UserClass(demo_users)
```

---

## 3. Creating and Managing Assignments

Assignments represent fieldwork tasks allocated to specific interviewers or supervisors.

### Creating Preloaded Assignments

Assignments can be created with pre-filled sample identifiers (e.g. household ID, cluster, address, or target respondent name) using `suso_createASS()`:

```r
# Create assignments for a questionnaire
suso_createASS(
  questID = "17a9fa22-1ef6-46d2-8c30-426aa876f273",
  version = 1,
  df = list(
    ResponsibleName = "interviewer01",
    Quantity = 10
  )
)
```

### Querying Existing Assignments with `assignmentClass`

`suso_get_assignments()` retrieves all assignments in the active workspace:

```r
# Retrieve all assignments
assignments <- suso_get_assignments()
```

The returned object is an `assignmentClass` instance:

```{r}
# View sample assignments
demo_ass[, .(Id, ResponsibleName, Quantity, InterviewsCount, Archived)]
```

Check class membership:

```{r}
is.assignmentClass(demo_ass)
```

### Reassigning Work in the Field

If an enumerator falls ill or is reassigned, transfer their assignment to another team member:

```r
# Reassign assignment #101 to interviewer02
suso_set_assignments(
  assignment_id = 101,
  reassign = TRUE,
  responsible = "interviewer02"
)
```

---

## 4. Map Management for Spatial Surveys

For spatial surveys, Survey Solutions supports offline basemaps (e.g. GeoTIFF, MBTiles, Shapefiles) downloaded to tablets.

### Uploading and Assigning Maps

```r
# Upload a new offline map
suso_mapupload(
  filePath = "maps/cluster_101_boundary.tif"
)

# Assign map to an interviewer
suso_mapassign(
  fileName = "cluster_101_boundary.tif",
  userName = "interviewer01",
  assignUser = TRUE
)
```

### Generating Spatial Reports

```r
# Generate a map report of completed interviews
suso_mapreport(
  questID = "17a9fa22-1ef6-46d2-8c30-426aa876f273",
  version = 1,
  variable = "gps_dwelling"
)
```
